Cognitive Context: Impact on the Perception of a Robot

### 2015-07-08 main.tex
Implemented Pierre's comments

We take here a different approach to measure anthropomorphism, based on
eye-tracking.  By looking at the eyes fixation on the head of the robot, we have
been able to demonstrate significantly different patterns in two experimental
conditions subtly different in terms of human-likeness. This eye-tracking based
metric is behavioural and hence does not suffer post-hoc reconstruction by the
subjects, possibly providing a better, ``noise-less'', access to the human
attitude towards the machine, and, as such, a better proxy to the cognitive and
affective bonds that the human establishes with the robot.

@@

The study has been designed out of the following intuition: \emph{the more we
ascribe cognitive capabilities to an agent, the more we focus on its head when
interacting with it}. Eye-tracking would be a natural methodology to investigate
this intuition, and we would then design audio-visual stimuli that would elicit different
levels of cognitive ascription onto the robot while being identical in every
other way.

@@

Finally, the hypothesis \h{4} examines whether the cognitive context affects the
general tendency to anthropomorphize, \ie whether observing \emph{one}
robot acting in an \emph{human-like} context leads to an increased level of
anthropomorphic attributions to robots \emph{in general}. This also provides us
with a way to triangulate the behavioural measures with the subjective
attribution of anthropomorphism: if such an effect is evidences, we can also
conclude that the cognitive context impacts the attribution of human-likeness,
and this, in turns, supports the hypothesis \h{3} that head fixations would
reflect not only the cognitive context but also the level of anthropomorphic
attributions.

@@

    \caption{\small The study follows a 2$\times$2 design: the \emph{robot} vs
        \emph{human} condition is within subject, while the \emph{shallow
        cognitive context} \vs \emph{deep cognitive context} is between subject.
        Two stimuli (\emph{Picking} task, {\bf A}, and \emph{Sound} task,
        {\bf B}) were shown to the participants, introduced by brief verbal
        commands, reproduced in the table. Figure~\ref{fig:stimuli} shows
        pictures of the four stimuli.}

@@

Four different video stimuli where filmed (figure~\ref{fig:stimuli}): two
different tasks (picking a stuffed animal, refered hereafter as the
\emph{Picking} task, and pointing towards a source of sound, refered as the
\emph{Sound} task) acted either by a human or a robot. Video stimuli were
realised in studio conditions. All videos followed the same simple structure: an
initial audio command (spoken by an invisible person), followed by the task
being executed.

The human actor was instructed to follow as closely as possible the actions and
attitudes of the robot (left/right glances, hesitations, gestures), while keeping a
natural, \emph{human-like} general behaviour.  Hence, the length of human videos
(57 seconds in total for the two tasks) was shorter than robot videos (110
seconds) due to the robot being usually slower in performing actions (in
particular walks) compared to human.

@@

This simple verbal priming (\emph{``brown toy''} \vs \emph{``your favorite
toy''}) set the cognitive context (\emph{shallow} \vs \emph{deep}) and was
expected to lead to different perceptions of the interaction
(\emph{machine-like} \vs \emph{human-like}). We reinforced this effect in the
\emph{Sound} task: not only the verbal priming but also the sound played by the
speakers during the task was altered: either a repetitive bip (\emph{shallow
cognitive context}) or the sound of a crying baby (\emph{deeper cognitive
context}).

@@

    \subfigure[\emph{Sound} task, robot condition]{
        \includegraphics[height=5cm]{stimulus-robot-noise}
    }
    \subfigure[\emph{Sound} task, human condition]{
        \includegraphics[height=5cm]{stimulus-human-noise-anon}
    }

@@

\begin{figure}
    \centering
    \subfigure[\emph{Picking} task, robot condition]
    {\includegraphics[height=5cm]{aoi-robot}\label{fig:aoi-toys}}
    \subfigure[\emph{Sound} task, human condition]
    {\includegraphics[height=5cm]{aoi-human}\label{fig:aoi-noise}}

@@

Based on ANOVA results, taking the 5 groups of AOIs {\sf head}, {\sf arms}, {\sf
hands}, {\sf torso} and {\sf legs} for the robot in both the shallow and deep
cognitive context conditions, the fixations on the head in the deep condition
were found to be significantly higher than that in the shallow condition
($F[1,54] = 4.2, p < .05$). For all other AOIs groups, the fixations across
shallow and deep conditions were almost the same. Figure~\ref{h3} shows the
distribution of proportion of time spent gazing on the 5 AOI groups for the two
conditions.

@@

Our first contribution extends our understanding of the attribution of
human-like characteristics to robot by exploring the role of the interaction
context. As we show, it appears that \textbf{even relatively
subtle context priming may lead to significantly different anthropomorphic
perceptions of the exact same robot, performing the exact same task}, as
confirmed both by distinct gaze patterns, and different reported perceptions in
questionnaires.

@@

We initially prepared three different stimuli: the \emph{Picking} task, the
\emph{Sound} task and a \emph{Dance} task.  During the \emph{Dance} task, the
agent was either asked to \emph{``Show some movements''} (\emph{shallow
cognitive context} condition) or to \emph{``Dance with the music''}
(\emph{deeper cognitive context} condition). While all the participants were
shown this stimuli as well, we decided to exclude it from the results as the
nature of the task (participants had to watch body movements) defeated our
gaze-based investigation method: the subjects expectedly paid attention to each
and every part of the body of the robot or human (even distribution of the gazes
on the body), thus making specific gaze patterns on the head difficult to
evidence.

@@

We would like here to acknowledge the contribution of Julia Fink to a better
understanding of anthropomorphism as a psychological phenomenon.

>>>>

We explore here a different approach to measure anthropomorphism, based on
eye-tracking.  By looking at the eyes fixation on the head of the robot, we have
been able to demonstrate significantly different patterns in two experimental
conditions subtly different in terms of human-likeness. This eye-tracking based
metric is behavioural and hence does not suffer post-hoc reconstruction by the
subjects, possibly providing a better, ``noise-less'', access to the human
attitude towards the machine, and, as such, a better proxy to the cognitive and
affective bonds that the human establishes with the robot.

@@

The study has been designed out of the following intuition: \emph{the more we
ascribe cognitive capabilities to an agent, the more we focus on its head when
interacting with it}. Eye-tracking is a natural methodology to investigate
this intuition. We hence design audio-visual stimuli that elicit different
levels of cognitive ascription onto the robot while being identical in every
other way.

@@

Finally, the hypothesis \h{4} examines whether the cognitive context affects the
general tendency to anthropomorphize, \ie whether observing \emph{one}
robot acting in an \emph{human-like} context leads to an increased level of
anthropomorphic attributions to robots \emph{in general}. This also provides us
with a way to triangulate the behavioural measures with the subjective
attribution of anthropomorphism: if such an effect is found, we can also
conclude that the cognitive context impacts on the attribution of human-likeness,
and this, in turns, supports the hypothesis \h{3} that head fixations would
reflect not only the cognitive context but also the level of anthropomorphic
attributions.

@@

    \caption{\small The study follows a 2$\times$2 design: the \emph{robot} vs
        \emph{human} condition is within subject, while the \emph{shallow
        cognitive context} \vs \emph{deep cognitive context} is between subject.
        Two stimuli (\emph{Picking} task, {\bf A}, and \emph{Pointing} task,
        {\bf B}) were shown to the participants, introduced by brief verbal
        commands, reproduced in the table. Figure~\ref{fig:stimuli} shows
        pictures of the four stimuli.}

@@

Four different video stimuli where filmed (figure~\ref{fig:stimuli}): two
different tasks (picking a stuffed animal, refered hereafter as the
\emph{Picking} task, and pointing towards a source of sound, refered as the
\emph{Pointing} task) acted either by a human or a robot. Video stimuli were
realised in studio conditions. All videos followed the same simple structure: an
initial audio command (spoken by an invisible person), followed by the task
being executed.

The human actor was instructed to follow as closely as possible the actions and
attitudes of the robot (left/right glances, hesitations, gestures), while keeping a
natural, \emph{human-like} general behaviour. As a result, the length of human videos
(57 seconds in total for the two tasks) was shorter than robot videos (110
seconds) due to the robot being usually slower in performing actions (in
particular walks) compared to human.

@@

This simple verbal priming (\emph{``brown toy''} \vs \emph{``your favorite
toy''}) set the cognitive context (\emph{shallow} \vs \emph{deep}) and was
expected to lead to different perceptions of the interaction
(\emph{machine-like} \vs \emph{human-like}). We reinforced this effect in the
\emph{Pointing} task: not only the verbal priming but also the sound played by the
speakers during the task was altered: either a repetitive bip (\emph{shallow
cognitive context}) or the sound of a crying baby (\emph{deeper cognitive
context}).

@@

    \subfigure[\emph{Pointing} task, robot condition]{
        \includegraphics[height=5cm]{stimulus-robot-noise}
    }
    \subfigure[\emph{Pointing} task, human condition]{
        \includegraphics[height=5cm]{stimulus-human-noise-anon}
    }

@@

\begin{figure}
    \centering
    \subfigure[\emph{Picking} task, robot condition]
    {\includegraphics[height=5cm]{aoi-robot}\label{fig:aoi-toys}}
    \subfigure[\emph{Pointing} task, human condition]
    {\includegraphics[height=5cm]{aoi-human}\label{fig:aoi-noise}}

@@

We consider the 5 groups of AOIs {\sf head}, {\sf arms}, {\sf hands}, {\sf
torso} and {\sf legs} (the left and right arms, hands and legs are grouped
together). Based on ANOVA results for the robot in both the shallow and deep
cognitive context conditions, the fixations on the head in the deep condition
were found to be significantly higher than that in the shallow condition
($F[1,54] = 4.2, p < .05$). For all other AOIs groups, the fixations across
shallow and deep conditions were almost the same. Figure~\ref{h3} shows the
distribution of proportion of time spent gazing on the 5 AOI groups for the two
conditions.

@@

Our first contribution extends our understanding of the attribution of
human-like characteristics to robot by exploring the role of the interaction
context. As we show, it appears that \textbf{even relatively subtle context
priming may lead to significantly different anthropomorphic perceptions of
the exact same robot, performing the exact same task}, as confirmed both by
distinct gaze patterns, and different reported perceptions in questionnaires.
Interestingly, this effects can already be evidenced after \textbf{very short,
one-way, interactions} (less than 2 minutes).

@@

We initially prepared three different stimuli: the \emph{Picking} task, the
\emph{Pointing} task and a \emph{Dance} task.  During the \emph{Dance} task, the
agent was either asked to \emph{``Show some movements''} (\emph{shallow
cognitive context} condition) or to \emph{``Dance with the music''}
(\emph{deeper cognitive context} condition). While all the participants were
shown this stimuli as well, we decided to exclude it from the results as we
found out that the nature of the task (participants had to watch body movements)
introduced an experimental bias: because dance is a form of whole-body
expression, gazes where evenly distributed on the body of the agents, masking
possible cognition-related gaze patterns.

@@

We would like here to acknowledge the contribution of Julia Fink to a better
understanding of anthropomorphism as a psychological phenomenon. We also thank
all the students for their participation.

This research was supported by the Swiss National Science Foundation through the
National Centre of Competence in Research Robotics.
